---
title: "Reproducible Course Project 1"
author: "Ophelia"
date: "June 14, 2016"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


## part1: read and process data
This is my first knitr document.  The actiity monitoring dataset is stored in a 
csv file and there are 17,568 observation and 3 variables (steps, date, interval) in this dataset.  
It is downloaded from 
https://www.coursera.org/learn/reproducible-research/peer/gYyPt/course-project-1
Here is code chunk reading data from working directory

```{r chunk_read}
dataset <- read.csv("activity.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)
```


## part2: histogram of total number steps taken per day
Here is code chunk generating plot

```{r chunk_histogram1}
aggData1 <- aggregate(dataset[,'steps'], by=list(Group.date=dataset$date),FUN=sum,na.rm=TRUE)
aggData1$Group.date <- as.Date(aggData1$Group.date)
head(aggData1)
hist(aggData1$x,col="blue",main="total steps per day",xlab="steps",freq=TRUE)
```
The aggregated data summarizing total number steps taken each day with missing data removed 
contains 61 observations, of which 8 of them are zeros steps.

## part3: mean and median values of part2
Here is code chunk of summary of total steps taken each day

```{r chunk_mean}
summary(aggData1)
```
The mean of total numbers of steps taken per day is 9354.
The median of total numbers of steps taken per day is 10395.

## part4: time series plot of average steps taken
Here is code chunk generating plot

```{r chunk_time_series}
library(ggplot2)
aggData2 <- aggregate(dataset[,'steps'], by=list(Group.interval=dataset$interval),FUN=mean,na.rm=TRUE)
head(aggData2)
ggplot(aggData2,aes(Group.interval,x))+geom_line()+xlab("intervals") +ylab("steps")+
ggtitle("averge number steps taken across all days")
```


## part5: 5-minute interval contains maximum number of steps 
Here is code chunk to evaluate average number of steps in 5-minute intervals 

```{r chunk_max}
maxStep <- aggData2[which.max(aggData2$x),]
```
Interval `r maxStep[1,1]` contains maximuum steps of `r as.integer(maxStep[1,2])`.

## part6: strategy for imputing missing data 
Here is code chunk checking which variables misses more than 5% of data, 
a safe maximum threshold for dataset

```{r chunk_missing}
pMis <- function(x)(sum(is.na(x)/length(x)*100))
pMiss <- apply(dataset,2,pMis)
pMiss
```
We see that variable steps consists 13% (2304) NAs of total 17568 rows, 
Here is code chunk using mice package to replace missing data with 
imputed values in the first of five dataset

```{r chunk_impute}
library(mice)
tempData1 <- mice(dataset,m=5,maxit=20,meth='pmm',seed=500)
completedData <- complete(tempData1,1,include=TRUE)
```


## part7: histogram of total number steps taken per day after missing values are imputed
Here is code chunk generating plot

```{r chunk_histogram2}
aggData3 <- aggregate(completedData[,'steps'], by=list(Group.date=dataset$date),FUN=sum,na.rm=TRUE)
aggData3$Group.date <- as.Date(aggData3$Group.date)
hist(aggData3$x,col="blue",main="total steps per day",xlab="steps",freq=TRUE)
summary(aggData3)
```
The mean of total numbers of steps taken per day with imputed values is 10761.
The median of total numbers of steps taken per day with imputed values is 10654.
The mice function draws the mean and median closer together, making the distribution
more symmetrical.


## part8: panel plot average number of steps taken per 5-minute intervals
## comparing weekends and weekdays, using imputed values
Here is code chunk generating plot

```{r chunk_panel_series}
temp <- completedData
head(temp) 
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
temp$date <- as.Date(temp$date)
temp$wDay <- c('weekend','weekday')[(weekdays(temp$date) %in% weekdays1)+1L]
head(temp)
aggData4 <- aggregate(steps ~ interval+wDay, temp, mean) 
ggplot(aggData4,aes(interval,steps))+geom_line()+xlab("intervals") +ylab("steps")+
facet_wrap(~wDay,ncol=1)
```

